---
title: "Coursera Statistical Inference Project"
author: "Jiajun Zhang"
date: "October 12, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1: Simulation Exercise
 
 - The purpose of this exercise is to investigate the exponential distribution and make a simulated distribution and compare it to the theoretical one. 

```{r}
#Given Information 
lambda <- 0.2; n <- 40
B <- 1000 #Thousand simulations
```


### Sample Mean and Theoretical Mean of the Distribution

  - Here, we will find out the averages of 40 exponentials and do a thousand simulations. Then calculate the average of 1000 means. 
```{r, message=FALSE, fig.height=3, fig.width=6}
set.seed(123) 
#Simulations
simu_data <- matrix(rexp(n*B, lambda), nrow=B, ncol=n)
#Calculate the mean of each simulation
simu_mean <- apply(simu_data, 1, mean) #1 indicates row calculation
hist(simu_mean, breaks=20, xlab="Means", cex.main=0.8,
     main="Means of 1000 Simulations of Exponential Distribution")
abline(v=1/lambda, col=2)
abline(v=mean(simu_mean), col=3)
legend("topright", c("Theoretical Mean", "Sample Mean"), 
       col=c(2,3), lty=c(1,1), cex=0.85)
```

```{r}
#Theoretical Mean
1/lambda

#Simulation Mean
mean(simu_mean)
```

 - We see that from the graph and calculation, the simulation mean and theoretical mean of the exponential distribution are really close to each other. 

&nbsp;  


### Sample Variance and Theoretical Variance of the Distribution

  - Similar to the previous one, but this time we will compare their variances. 
```{r, fig.height=4, fig.width=6}
# Calculate the variances of each simulation
simu_var <- apply(simu_data, 1, var)
simu_var <- simu_var/n
hist(simu_var, breaks=20, xlab="Variances", xlim=c(0, 2), cex.main=0.8,
     main="Variances of 1000 Simulations of Exponential Distribution")
abline(v=((1/lambda)^2)/n, col=2)
abline(v=mean(simu_var), col=3)
legend("topright", c("Theoretical Variance", "Sample Variance"), 
       col=c(2,3), lty=c(1,1), cex=0.85)
```

```{r}
#Theoretical Variance
((1/lambda)^2)/n

#Simulation Variance
mean(simu_var)
```

 - We see that from the graph and calculation, the simulation variance and theoretical variance of the exponential distribution are really close to each other. 


&nbsp;  


### Distribution Normality

 - Now, we want to determine whether the simulation means follow a normal distribution. 
```{r, fig.height=4, fig.width=6}
#Plot a normal curve
x <- seq(min(simu_mean), max(simu_mean), length=200)
y <- dnorm(x, mean=1/lambda, sd=(1/lambda)/sqrt(n))
hist(simu_mean, breaks=20, prob=T, xlab="Means", 
     main="Means of 1000 Simulations of Exponential Distribution 
     \n and Normal Distribution Curve", cex.main=0.8)
lines(x, y, col=2)
```
 
 - The histogram shows the distribution of the means of 1000 simulated results and we add a normal distribution curve on it. The graph tells us that the simulated distribution approximately follows a normal shape. 
 
 
 


